---
title: "p8105_hw6_yh3683"
author: "Yining He"
date: "2024-11-21"
output: github_document
---
```{r include=FALSE}
library(tidyverse)
library(rnoaa)
library(dplyr)
library(broom)
library(purrr)
library(ggplot2)
library(modelr)
library(mgcv)
library(readxl)
```


# Problem1
```{r echo=FALSE}
weather_df <- rnoaa::meteo_pull_monitors(
  c("USW00094728"),
  var = c("PRCP", "TMIN", "TMAX"),
  date_min = "2017-01-01",
  date_max = "2017-12-31"
) %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10
  ) %>%
  select(name, id, everything())

view(weather_df)
```

```{r echo=FALSE}
# Set seed for reproducibility
set.seed(123)

# Generate 5000 bootstrap samples and compute statistics
bootstrap_results = tibble(strap_number = 1:5000) %>%
  mutate(
    # Generate bootstrap samples
    strap_sample = map(strap_number, ~sample_n(weather_df, size = nrow(weather_df), replace = TRUE)),
    
    # Fit models for each bootstrap sample
    models = map(strap_sample, ~lm(tmax ~ tmin, data = .x)),
    
    # Extract R-squared
    r_squared = map_dbl(models, ~glance(.x)$r.squared),
    
    # Extract coefficients and compute log(β0*β1)
    log_betas = map_dbl(models, ~{
      coefs = tidy(.x)$estimate
      log(coefs[1] * coefs[2])
    })
  )
```

```{r echo=FALSE}
# Create visualizations
# Distribution of R-squared
p1 = ggplot(bootstrap_results, aes(x = r_squared)) +
  geom_density(fill = "lightblue", alpha = 0.5) +
  labs(
    title = "Bootstrap Distribution of R-squared",
    x = "R-squared",
    y = "Density"
  ) +
  theme_minimal()

# Distribution of log(β0*β1)
p2 = ggplot(bootstrap_results, aes(x = log_betas)) +
  geom_density(fill = "lightgreen", alpha = 0.5) +
  labs(
    title = "Bootstrap Distribution of log(β0*β1)",
    x = "log(β0*β1)",
    y = "Density"
  ) +
  theme_minimal()

# Display both plots side by side
library(patchwork)
p1 + p2
```
```{r echo=FALSE}
# Calculate 95% confidence intervals
r_squared_ci = quantile(bootstrap_results$r_squared, c(0.025, 0.975))
log_betas_ci = quantile(bootstrap_results$log_betas, c(0.025, 0.975))
```
```{r}
cat("\n95% Confidence Interval for R-squared:", 
    "\nLower bound:", round(r_squared_ci[1], 4),
    "\nUpper bound:", round(r_squared_ci[2], 4))
```

The bootstrap analysis with 5000 samples of the 2017 Central Park weather data shows R-squared values ranging from 0.89 to 0.93 (95% CI), with a symmetrical distribution centered around 0.91. This indicates that minimum temperature explains approximately 90% of the variation in maximum temperature, demonstrating a remarkably strong predictive relationship.

```{r}
cat("\n\n95% Confidence Interval for log(β0*β1):", 
    "\nLower bound:", round(log_betas_ci[1], 4),
    "\nUpper bound:", round(log_betas_ci[2], 4))
```
The log(β0*β1) estimates display a normal distribution centered around 2.0, with a 95% confidence interval of [1.96, 2.06]. This entirely positive and narrow interval confirms a significant and stable positive relationship between minimum and maximum temperatures, providing strong evidence for the reliability of our linear model.
